%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Assumptions}
\label{sec:assumptions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We make a few assumptions concerning the environment or the
distributed system whose traces we analyze.

\begin{itemize}

\item We assume a non-broadcast networking medium at the message
  capture layer. For example, we do not handle wireless environments
  in which the nodes in the system sniff and thereby receive all
  messages. However, we can operate in a wireless environment with all
  nodes restricted to using TCP for communication, or in which all
  sniffed traffic is also logged in the trace.

\item We assume that the nodes are using a reliable communication
  protocol so that messages are never lost. This assumption may be
  dropped if in addition to message transmission the trace also
  includes message reception. In the absence of observable reception,
  Synoptic cannot distinguish between messaging behavior at the
  receiver node caused by a lost message and receiver behavior caused
  by message reception.

\item We assume that messages reordering does not occur. This is
  necessary to prevent re-orderings from adversely impacting our
  temporal invariant inference algorithm. A more robust algorithm can
  relax this assumption to a certain degree.

  % \item We assume that message transmission time is non-zero.

% \item We assume that the set of nodes in the system is constant and
%   does not change throughout the trace. That is, nodes do not fail and
%   new nodes do not join the system.

\item If the system is composed of just one physical host the local
  clock can be used for timestamping all messages. To analyze the
  behavior of multiple network nodes, a consistent notion of time is
  necessary. As a simplification, we assume correctly synchronized
  clocks (e.g. nodes use NTP). This assumption can be relaxed by
  establishing a global partial ordering on all the messages in the
  system with a vector clock algorithm~\cite{LamportClocks}.

% \item To simplify message capture and processing Synoptic assumes that
%   the system uses protobufs~\cite{protobuf}. However, this is not a
%   fundamental assumption, and is an artifact of our implementation.

\end{itemize}