%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this paper we presented Synoptic, a tool to summarize message
activity in a distributed system and to output a finite state automata
representation of this activity. Synoptic supports two algorithms --
Bikon and GK-Tail, which refine and coarsen the graph representation
respectively. We presented the technical details behind these
algorithms, and the ways in which Synoptic abstracts away the policy
to guide these algorithms into what we term the decision engine.  Our
results indicate that Synoptic produces representations for real
distributed systems traces, which can aid in understanding complex
system behaviors. In addition, we carried out a user study which
suggests that Synoptic-produced representations can verify the
intuition of a distributed systems developer. Synoptic is an open
source tool, which can be found at the following URL:
\texttt{\footnotesize
  http://code.google.com/p/cse503-system-analysis/}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Invaluable feedback on earlier drafts of this report by Michael Ernst
were instrumental in shaping our presentation and ideas into their
present form. We also thank him for making copious time to meet with
us on demand throughout the quarter. We also gratefully acknowledge
Ethan Katz-Bassett for his time to explain his reverse traceroute
system and for his participation in our user study.
